Relational programming in Smalltalk

### 2018-12-07 booklet/article.tex
Little more introduction paragraphs.

\begin{abstract}
This booklet is an extended description of $mu$Kanren, a relational interpreter
formerly implemented in Scheme and ported in Smalltalk for the sake of
understanding and, of course, fun. It enjoy simplicity and elegance despite the
complex nature of logic systems; moreover, using an uniform but powerful
language such as Smalltalk it is possible to appreciate and benefit from them
all.
\end{abstract}

@@

I believe \textit{microkanren} is, first of all, an \textit{educational beast},
concerning unification, lazy streams, backtracking and optimization; the
abstract definition was shown in \citep{Hemann:muKanren}.
I repeat the exercise of writing it in \citep{Nocentini:kanrens}:
\begin{itemize}
\item Python, native \verb|generator|s :) limits on recursive calls :(
\item OCaml, algebraic datatypes :) hard to extend :(
\item Smalltalk, simple, fast and clear :) many dispatching msgs :/
\end{itemize}

>>>>

\begin{abstract}
This booklet is an extended description of $\mu$Kanren, a relational
interpreter formerly implemented in Scheme and ported in Smalltalk for the sake
of understanding and, of course, fun. It enjoy simplicity and elegance despite
the complex nature of logic systems; moreover, using an uniform but powerful
language such as Smalltalk it is possible to appreciate and benefit from both
of them.
\end{abstract}

@@

This document extends the presentation of $\mu$Kanren shown at the last ESUG
conference \citep{ESUG2018} in two ways: (i)~it recaps existing ``literature''
on the subject and (ii)~it describes in more details our implementation coded
in Smalltalk.

It is a waste of time to repeat ourself on $\mu$Kanren's definition; on the
contrary, we believe interesting to look up and to watch over the many
implementation it has inspired since its seminal paper \citep{Hemann:muKanren}
and how programmers stretch the canonical implementation to take advantage of
their favorite programming language's features.

Certainly the book \citep{Friedman:Reasoned:Schemer} is the main reference
written in the same style of the other \emph{little books} of which Dan
Friedman coauthors them all; moreover, Byrd's dissertation \citep{Byrd:PhD}
shows advanced stuff and techniques while the repository
\citep{byrd:minikanren:repo} -- updated regularly by Byrd himself -- keeps
pointers to implementations, articles and presentations. All these references
uses the Scheme programming language, $\mu$Kanren is \emph{purely functional}
therefore any language that has the same power of the $\lambda$-calculus is a
candidate for implementing it, Scheme in particular; in parallel,
\citep{byrd:minikanren} has the same nature but uses Scheme macrology and in
some aspect it requires little effort to port to other languages that haven't
macros.

I believe \textit{microkanren} is, first of all, an \textit{educational beast},
concerning unification, lazy streams, backtracking and optimization; the
abstract definition was shown in 
I repeat the exercise of writing it in \citep{Nocentini:kanrens}:
\begin{itemize}
\item Python, native \verb|generator|s :) limits on recursive calls :(
\item OCaml, algebraic datatypes :) hard to extend :(
\item Smalltalk, simple, fast and clear :) many dispatching msgs :/
\end{itemize}
